---
title: "analysis"
author: "pc2853"
date: "November 23, 2018"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(lubridate)
library(xml2)
library(rvest)
library(patchwork)
library(ggplot2)
library(knitr)
```

This analysis is based on the final data set "finaldat.RData" and its subset data corresponding to 2 neighborhoods "subset_dat.RData". These two datasets are built by running the code from "Data_prep.Rmd" and "Data_merge.Rmd" by Dew. The default significance level is 0.05.

```{r}
load('data/finaldat.RData')


load('data/subset_dat.RData')


```

1. hour of the day vs log response time ---- ANOVA MODEL

```{r}
fit_1 = lm(log_response_time~hour_of_day, data = finaldat)
summary(fit_1)
kable(anova(fit_1))
```

Ho: different hour of the day makes no difference in logs response time.
p-value: very small.
conclusion: There is a difference of log response time in different time/hour of the day.

2. snow & prcp vs log response time ----- linear regression
```{r}
fit_2 = lm(log_response_time~snow, data = finaldat)
fit_2 %>% 
  broom::tidy() %>% 
  knitr::kable(digits = 3)
fit_2_2 = lm(log_response_time~prcp, data = finaldat)
fit_2_2 %>% 
  broom::tidy() %>% 
  knitr::kable(digits = 3)

```


H0: There is no correlation between snow and log_response_time
p-value: very small
estimate: positive
conclusion: There is a positive correlation between snow and log response time. 

3. seasons vs log response time ----- ANOVA
```{r}
fit_3 = lm(log_response_time~season, data = finaldat)
kable(anova(fit_3))
```

Ho: different seasons make no difference in logs response time.
p-value: very small.
conclusion: There is a difference of log response time in different seasons.

4. MLR based on snow, prcp, tmax, tmin(factors of nature environment) in different hour of the day----- MLR model
```{r}
fit_4 = lm(log_response_time~snow + prcp + tmax + tmin, data = finaldat)
fit_4 %>% 
  broom::tidy() %>% 
  knitr::kable(digits = 3)

fit_4_2 = finaldat %>% 
  group_by(hour_of_day) %>% 
  nest() %>% 
  mutate(models = map(data, ~lm(log_response_time ~ snow + prcp + tmax + tmin, data = .x)),
         models = map(models, broom::tidy)) %>% 
  select(-data) %>% 
  unnest()

summary(fit_4_2)
fit_4_2 %>% 
  select(hour_of_day, term, estimate) %>% 
  mutate(term = fct_inorder(term)) %>% 
  spread(key = term, value = estimate) %>% 
  knitr::kable(digits = 3)
```

Despite there are correlations, the estimates are really small. 

## Logistic regression with binary outcome 5min- / 5min+
```{r}
## Added snow and prcp categories
finaldat_ctg = 
  finaldat %>% 
  mutate(over_5min = if_else((over_5min == "5min-"), 0, 1),
         prcp_ctg = 
           if_else(prcp == 0, "no_prcp",
           if_else((prcp > 0 & prcp <= 25), "low", "high")),
         snow_ctg = 
           if_else(snow == 0, "no_snow",
           if_else((snow > 0 & snow <= 50), "low", "high")),
         prcp_ctg = fct_relevel(prcp_ctg, "no_prcp"),
         snow_ctg = fct_relevel(snow_ctg, "no_snow"),
         season = fct_relevel(season, "Spring"),
         hour_of_day = fct_relevel(hour_of_day, "morning")) 

fit_logistic =
  glm(over_5min ~ season + hour_of_day + snow_ctg + prcp_ctg, 
      family = binomial(), data = finaldat_ctg)

summary(fit_logistic)

fit_logistic %>% 
  broom::tidy() %>% 
  mutate(OR = exp(estimate),
         lower_CI_OR = exp(estimate - 1.96*std.error),
         upper_CI_OR = exp(estimate + 1.96*std.error)) %>% 
  select(term, OR, lower_CI_OR, upper_CI_OR) %>% 
  mutate(term = fct_reorder(term, OR)) %>% 
  ggplot(aes(x = term, y = OR)) + 
  geom_point() + 
  geom_errorbar(aes(ymin = lower_CI_OR, ymax = upper_CI_OR), 
                color = "darkred") +
  coord_flip() +
  geom_hline(yintercept = 1.0, color = "blue") +
  theme_bw()

## prcp_ctg
fit_logistic %>% 
  broom::tidy() %>% 
  filter(str_detect(term, "prcp_ctg")) %>% 
  mutate(OR = exp(estimate),
         lower_CI_OR = exp(estimate - 1.96*std.error),
         upper_CI_OR = exp(estimate + 1.96*std.error)) %>% 
  select(term, OR, lower_CI_OR, upper_CI_OR) %>% 
  mutate(term = fct_reorder(term, OR)) %>% 
  ggplot(aes(x = term, y = OR)) + 
  geom_point() + 
  geom_errorbar(aes(ymin = lower_CI_OR, ymax = upper_CI_OR), 
                color = "darkred") +
  scale_y_continuous(limits = c(0.5, 1.8)) +
  coord_flip() +
  geom_hline(yintercept = 1.0, color = "blue") +
  theme_bw() 

## snow_ctg
fit_logistic %>% 
  broom::tidy() %>% 
  filter(str_detect(term, "snow_ctg")) %>% 
  mutate(OR = exp(estimate),
         lower_CI_OR = exp(estimate - 1.96*std.error),
         upper_CI_OR = exp(estimate + 1.96*std.error)) %>% 
  select(term, OR, lower_CI_OR, upper_CI_OR) %>% 
  mutate(term = fct_reorder(term, OR)) %>% 
  ggplot(aes(x = term, y = OR)) + 
  geom_point() + 
  geom_errorbar(aes(ymin = lower_CI_OR, ymax = upper_CI_OR), 
                color = "darkred") +
  scale_y_continuous(limits = c(0.5, 1.8)) +
  coord_flip() +
  geom_hline(yintercept = 1.0, color = "blue") +
  theme_bw()

## Hour of day
fit_logistic %>% 
  broom::tidy() %>% 
  filter(str_detect(term, "hour_of_day")) %>% 
  mutate(OR = exp(estimate),
         lower_CI_OR = exp(estimate - 1.96*std.error),
         upper_CI_OR = exp(estimate + 1.96*std.error)) %>% 
  select(term, OR, lower_CI_OR, upper_CI_OR) %>% 
  mutate(term = fct_reorder(term, OR)) %>% 
  ggplot(aes(x = term, y = OR)) + 
  geom_point() + 
  geom_errorbar(aes(ymin = lower_CI_OR, ymax = upper_CI_OR), 
                color = "darkred") +
  coord_flip() +
  geom_hline(yintercept = 1.0, color = "blue") +
  theme_bw()
```

## (1) Mean response time in snowy condition

```{r}
## OR result above matches with the results below
finaldat %>% 
  mutate(snow = snow > 0) %>%
  group_by(date, snow) %>% 
  summarise(mean_resp_time = mean(response_time)) %>% 
  ggplot(aes(x = snow, y = mean_resp_time)) +
  geom_violin(aes(fill = factor(snow)), alpha = .5) +
  stat_summary(fun.y = mean, geom = "point", size = 4, color = "blue") +
  labs(
    title = "Mean response time in snowy conditions",
    y = "Mean response time",
    x = "Snow"
  ) +
  viridis::scale_fill_viridis(
    name = "Snow",
    discrete = TRUE) +
  theme(plot.title = element_text(size = 12),
        strip.background = element_rect(fill = "black"),
        strip.text = element_text(color = "white", face = "bold"),
        legend.position = "None") 
```

## (2) Response time (smoothed line) by hour of the day

```{r}
finaldat %>%
    group_by(hour) %>% 
  ggplot(aes(x = hour, y = response_time)) + 
  geom_smooth(se = FALSE) +
  labs(x = "Hour of the day", y = "Mean response time") +
  geom_vline(xintercept = c(6, 13, 18), color = "green") +
  theme_bw()
```

## Logistic regression with binary outcome 8min- / 8min+
```{r}

finaldat_8min =
  finaldat %>% 
  mutate(over_8min = ifelse(response_time > 8, "8min+", "8min-"))
## Added snow and prcp categories
finaldat_ctg_8min = 
  finaldat_8min %>% 
  mutate(over_8min = if_else((over_8min == "8min-"), 0, 1),
         prcp_ctg = 
           if_else(prcp == 0, "no_prcp",
           if_else((prcp > 0 & prcp <= 25), "low", "high")),
         snow_ctg = 
           if_else(snow == 0, "no_snow",
           if_else((snow > 0 & snow <= 50), "low", "high")),
         prcp_ctg = fct_relevel(prcp_ctg, "no_prcp"),
         snow_ctg = fct_relevel(snow_ctg, "no_snow"),
         season = fct_relevel(season, "Spring"),
         hour_of_day = fct_relevel(hour_of_day, "morning")) 

fit_logistic_8min =
  glm(over_8min ~ season + hour_of_day + snow_ctg + prcp_ctg, 
      family = binomial(), data = finaldat_ctg_8min)

summary(fit_logistic_8min)

fit_logistic_8min %>% 
  broom::tidy() %>% 
  mutate(OR = exp(estimate),
         lower_CI_OR = exp(estimate - 1.96*std.error),
         upper_CI_OR = exp(estimate + 1.96*std.error)) %>% 
  select(term, OR, lower_CI_OR, upper_CI_OR) %>% 
  mutate(term = fct_reorder(term, OR)) %>% 
  ggplot(aes(x = term, y = OR)) + 
  geom_point() + 
  geom_errorbar(aes(ymin = lower_CI_OR, ymax = upper_CI_OR), 
                color = "darkred") +
  coord_flip() +
  geom_hline(yintercept = 1.0, color = "blue") +
  theme_bw()

## prcp_ctg
fit_logistic_8min %>% 
  broom::tidy() %>% 
  filter(str_detect(term, "prcp_ctg")) %>% 
  mutate(OR = exp(estimate),
         lower_CI_OR = exp(estimate - 1.96*std.error),
         upper_CI_OR = exp(estimate + 1.96*std.error)) %>% 
  select(term, OR, lower_CI_OR, upper_CI_OR) %>% 
  mutate(term = fct_reorder(term, OR)) %>% 
  ggplot(aes(x = term, y = OR)) + 
  geom_point() + 
  geom_errorbar(aes(ymin = lower_CI_OR, ymax = upper_CI_OR), 
                color = "darkred") +
  scale_y_continuous(limits = c(0.5, 1.8)) +
  coord_flip() +
  geom_hline(yintercept = 1.0, color = "blue") +
  theme_bw() 

## snow_ctg
fit_logistic_8min %>% 
  broom::tidy() %>% 
  filter(str_detect(term, "snow_ctg")) %>% 
  mutate(OR = exp(estimate),
         lower_CI_OR = exp(estimate - 1.96*std.error),
         upper_CI_OR = exp(estimate + 1.96*std.error)) %>% 
  select(term, OR, lower_CI_OR, upper_CI_OR) %>% 
  mutate(term = fct_reorder(term, OR)) %>% 
  ggplot(aes(x = term, y = OR)) + 
  geom_point() + 
  geom_errorbar(aes(ymin = lower_CI_OR, ymax = upper_CI_OR), 
                color = "darkred") +
  scale_y_continuous(limits = c(0.5, 1.8)) +
  coord_flip() +
  geom_hline(yintercept = 1.0, color = "blue") +
  theme_bw()

## Hour of day
fit_logistic_8min %>% 
  broom::tidy() %>% 
  filter(str_detect(term, "hour_of_day")) %>% 
  mutate(OR = exp(estimate),
         lower_CI_OR = exp(estimate - 1.96*std.error),
         upper_CI_OR = exp(estimate + 1.96*std.error)) %>% 
  select(term, OR, lower_CI_OR, upper_CI_OR) %>% 
  mutate(term = fct_reorder(term, OR)) %>% 
  ggplot(aes(x = term, y = OR)) + 
  geom_point() + 
  geom_errorbar(aes(ymin = lower_CI_OR, ymax = upper_CI_OR), 
                color = "darkred") +
  coord_flip() +
  geom_hline(yintercept = 1.0, color = "blue") +
  theme_bw()
```


### estimate the effect of Street closure on response time in the selected two neighborhoods. 
5. street closure(binary data 0 or 1) vs log response time ------ t-test
```{r}
t.test(log_response_time~street_closed, data = dat_subset) %>% 
  broom::tidy() %>% 
  select(estimate, p.value, conf.low, conf.high, method, alternative) %>% 
  knitr::kable(diigts = 3)
#just curious about the visualization
dat_subset %>% 
  mutate(street_closed = as.factor(street_closed)) %>% 
  ggplot(aes(x = street_closed, y = log_response_time, color = neighborhood)) + 
  geom_boxplot()
#actually the difference is really small.  

```

H0: The presence of street closure affects log response time.
p value: 0.036 < 0.05
conclusion: Having street closure while incident happens affects the log response time. 

