---
title: "incident_EDA"
author: "Team"
date: "November 10, 2018"
output: github_document
---

```{r}
library(tidyverse)
library(lubridate)
library(xml2)
library(rvest)
library(patchwork)
```

##### Data sets ######

## Response time data - Loaded and tidied data
```{r include = FALSE}
incident_dat_2017 <-  
  read_csv('data/Incidents_Responded_to_by_Fire_Companies.csv',
  col_types = "cicccicciccccccccccciccccccc") %>% 
  janitor::clean_names() %>% 
  #recode date/time
  mutate(incident_date_time = mdy_hms(incident_date_time),
         arrival_date_time = mdy_hms(arrival_date_time)) %>% 
  #select year 2017
  filter(year(incident_date_time) == 2017,
         incident_type_desc == "300 - Rescue, EMS incident, other") %>% 
  select(im_incident_key, incident_date_time, arrival_date_time,
         street_highway:borough_desc) %>% 
  na.omit() %>% 

  # Added response time(minute) variable
  # mutate(response_time = arrival_date_time - incident_date_time) %>%
  mutate(response_time = difftime(arrival_date_time, incident_date_time, units = 'mins'),
  # Added hour variable
  hour = hour(incident_date_time),
  date = date(incident_date_time),
  # Added incident_month and incident_day variables from incident_date_time
  incident_date = as.Date(incident_date_time)) %>% 
  separate(incident_date, 
           into = c("incident_year", "incident_month", "incident_day"), 
           sep = "-") %>% 
  select(-incident_year) %>% 
  mutate(incident_month = as.numeric(incident_month),
         incident_day = as.numeric(incident_day))
```


## Added neighborhood variable
```{r fig.height = 8, include = FALSE}
url = "https://www.health.ny.gov/statistics/cancer/registry/appendix/neighborhoods.htm?fbclid=IwAR3N4VlKC1OehRZyEuDYPEAE7AFAEXXIRC11seIBKxA-0fd3g4hL0QvnV20"

xml = read_html(url)

zip_code_table = (xml %>% html_nodes(css = "table")) %>% 
  .[[1]] %>%
  html_table() %>% 
  janitor::clean_names() %>%  
  select(neighborhood, zip_codes) %>% 
  separate(zip_codes, c("a", "b", "c", "d", "e", "f", "g", "h", "i"), 
           sep = ",") %>% 
  gather(key = to_remove, value = zip_code, a:i) %>% 
  select(-to_remove) %>% 
  na.omit() %>% 
  distinct() %>% 
  mutate(zip_code = as.numeric(zip_code))

# To match zip_code to neighborhood table, converted it to numeric
incident_dat_2017 = 
  incident_dat_2017 %>% 
    mutate(zip_code = as.numeric(zip_code)) 

incident_dat_2017 =  
  left_join(incident_dat_2017, zip_code_table, by = "zip_code") 
```

## Add weather-related variable
```{r include = FALSE}
library(rnoaa)

nyc_weather_2017 = 
  rnoaa::meteo_pull_monitors("USW00094728", 
                             var = c("PRCP", "TMIN", "TMAX", "SNOW", "SNWD"), 
                             date_min = "2017-01-01", 
                             date_max = "2017-12-31"
                             ) %>% 
  mutate(tmin = tmin/10, tmax = tmax/10, prcp = prcp/10) %>% 
  select(-id)

incident_weather = 
  left_join(incident_dat_2017, nyc_weather_2017, by = "date")
```

## Added prcp category variable

```{r include = FALSE}
nyc_weather_2017 = 
  nyc_weather_2017 %>%
  mutate(prcp_ctg = 
           ifelse(prcp == 0, "NA",
           ifelse((prcp > 0 & prcp <= 25), "low",
           ifelse((prcp > 25 & prcp <= 60), "medium", "high"))))

incident_prcp_ctg = 
  inner_join(incident_dat_2017, nyc_weather_2017, by = "date")

incident_prcp_ctg %>%
  group_by(date) %>% 
  filter(prcp > 0, 
         neighborhood == "Kingsbridge and Riverdale") %>%
  ggplot(aes(x = prcp_ctg, y = response_time)) +
  geom_violin(aes(fill = factor(prcp_ctg)), alpha = .5) +
  stat_summary(fun.y = mean, geom = "point", size = 4, color = "blue")
```

## Added season variable
```{r}
incident_season = 
  incident_dat_2017 %>%
  mutate(season = 
           ifelse(incident_month %in% 9:11, "Fall",
           ifelse(incident_month %in% c(12,1,2), "Winter",
           ifelse(incident_month %in% 3:5, "Spring", "Summer"))))
```

##Added incident response time category variable 
```{r}
incident_by_date = 
  incident_dat_2017 %>% 
  group_by(zip_code, date) %>% 
  summarize(n_incident = n())  
  
incident_by_zipcode = 
  incident_dat_2017 %>% 
  group_by(zip_code, date) %>% 
  summarize(n_incident = n()) %>% 
  ungroup() %>% 
  group_by(zip_code) %>% 
  summarize(mean_incident = mean(n_incident),
            sd_incident = sd(n_incident),
            std_high_incident = mean_incident + sd_incident,
            std_low_incident = mean_incident - sd_incident)

incident_response_time =
  incident_dat_2017 %>% 
  group_by(zip_code) %>% 
  select(zip_code, response_time, date) 

incident_category_data = 
  left_join(incident_by_date, incident_by_zipcode, by = "zip_code") %>% 
  mutate(incident_category = 
           ifelse(((n_incident - std_high_incident) > 0), "high", 
           ifelse(((std_low_incident - n_incident) > 0), "low", "medium"))) %>% 
  select(zip_code, date, incident_category)

incid_cate_joined_data = 
  inner_join(incident_dat_2017, incident_category_data) %>% 
  na.omit() 
```

## added hour of the day variable
```{r}
incident_hour_of_day = 
  incident_dat_2017 %>% 
  mutate(hour_of_day = 
           ifelse(hour %in% 6:12, "morning",
           ifelse(hour %in% 13:17, "afternoon",
           ifelse(hour %in% 18:23, "night","dawn"))))
```

## Street closure data 
```{r}
street_closure_2017 <-  
  read_csv('data/Street_Closures_due_to_construction_activities_by_Intersection.csv') %>% 
  janitor::clean_names() %>% 
  #recode date/time
  mutate(work_start_date = mdy_hms(work_start_date),
         work_end_date = mdy_hms(work_end_date),
         work_time = round(difftime(work_end_date, work_start_date, 
                                    units = 'days'), 0)) %>% 
  #select year 2017
  filter(year(work_start_date) == 2017) %>% 
  select(-purpose) %>% 
  na.omit() 
```

###### EDA ######
## Response time by area(neighborhood)
```{r fig.height = 8, include = FALSE}
incident_dat_2017 %>%
  na.omit(neighborhood) %>% 
  group_by(neighborhood) %>%
  summarise(mean_response_time = mean(response_time, na.rm = TRUE)) %>% 
  mutate(neighborhood = forcats::fct_reorder(neighborhood, mean_response_time, 
                                         .asc = TRUE)) %>% 
  ggplot(aes(x = neighborhood, y = mean_response_time)) + 
  geom_point(size = 1, color = "darkred") +  
  coord_flip() +
  theme_bw()
```


## create frequency by day for the whole year
```{r}
incident_dat_2017 %>% 
  mutate(date = date(incident_date_time)) %>%
  group_by(date) %>% count() %>%
  ggplot(aes(x = date, y = n)) + geom_line() + labs(y = 'Frequency')
```

## Look at monthly trend in EMS rescue incident
```{r}
incident_dat_2017 %>% 
  group_by(incident_month) %>% count() %>% 
  ggplot(aes(x = incident_month, y = n)) + geom_line() + 
  theme(axis.text.x = element_text(angle = 45)) +  labs(y = 'Frequency') +
  scale_x_continuous(breaks = 1:12, labels = month.name)
```

## Look at hourly trend
```{r}
incident_dat_2017 %>% 
  group_by(hour) %>% 
  count() %>% 
  ggplot(aes(x = hour, y = n)) + geom_line() + labs(y = 'Frequency')
```

#And look at lag time (average) by the hour
```{r}
incident_dat_2017 %>%
    group_by(hour) %>% 
    summarise(mean_response_time = mean(response_time, na.rm = TRUE)) %>% 
  ggplot(aes(x = hour, y = mean_response_time)) + geom_line() + 
  labs(y = 'Mean response time (minutes)')
```

```{r}
incident_dat_2017 %>%
    group_by(hour) %>% 
  ggplot(aes(x = hour, y = response_time)) + 
  geom_smooth(se = FALSE) 
```

## Hour of the day vs frequency, mean response time
```{r}
library(grid)

hourly_freq = 
  incident_dat_2017 %>% 
  group_by(hour) %>% 
  count() %>% 
  ggplot(aes(x = hour, y = n)) + 
  geom_line() +
  labs(x = "Hour of the day", y = "Frequency") +
  theme_bw()

hourly_time = 
  incident_dat_2017 %>%
    group_by(hour) %>% 
  ggplot(aes(x = hour, y = response_time)) + 
  geom_smooth(se = FALSE) +
  labs(x = "Hour of the day", y = "Mean response time") +
  theme_bw()

grid.newpage()
grid.draw(rbind(ggplotGrob(hourly_freq), ggplotGrob(hourly_time), size = "last"))
```


## Month vs frequency, mean response time
```{r}
# month_freq =
#   incident_dat_2017 %>% 
#   group_by(incident_month) %>% count() %>% 
#   ggplot(aes(x = incident_month, y = n)) + geom_line() + 
#   theme(axis.text.x = element_text(angle = 45)) +  labs(y = 'Frequency') +
#   scale_x_continuous(breaks = 1:12, labels = month.name) +
#   theme_bw()
# 
# month_time = 
#   incident_dat_2017 %>%
#   group_by(incident_month) %>% 
#   ggplot(aes(x = incident_month, y = response_time)) + 
#   geom_smooth(se = FALSE) +
#   labs(x = "Month", y = "Mean response time") +
#   scale_x_continuous(breaks = 1:12, labels = month.name) +
#   theme_bw() +
#   theme(axis.text.x = element_text(angle = 45)) 
# 
# grid.newpage()
# grid.draw(rbind(ggplotGrob(month_freq), ggplotGrob(month_time), size = "last"))
```


### rainy day vs snowy day
```{r}
resp_time_prcp =
  incident_weather %>% 
  mutate(prcp = prcp > 0) %>%
  group_by(date, prcp) %>% 
  summarise(mean_resp_time = mean(response_time)) %>% 
  ggplot(aes(x = prcp, y = mean_resp_time)) +
  geom_violin(aes(fill = factor(prcp)), alpha = .5) +
  stat_summary(fun.y = mean, geom = "point", size = 4, color = "blue") +
  labs(
    title = "Mean Response time in rainy conditions",
    x = "Precipitation",
    y = "Response time"
  ) +
  viridis::scale_fill_viridis(
    name = "Precipitation",
    discrete = TRUE) +
  theme(plot.title = element_text(size = 12),
        strip.background = element_rect(fill = "black"),
        strip.text = element_text(color = "white", face = "bold"),
        legend.position = "None") 

resp_time_snow =
  incident_weather %>% 
  mutate(snow = snow > 0) %>%
  group_by(date, snow) %>% 
  summarise(mean_resp_time = mean(response_time)) %>% 
  ggplot(aes(x = snow, y = mean_resp_time)) +
  geom_violin(aes(fill = factor(snow)), alpha = .5) +
  stat_summary(fun.y = mean, geom = "point", size = 4, color = "blue") +
  labs(
    title = "Mean response time in snowy conditions",
    y = "Mean response time",
    x = "Snow"
  ) +
  viridis::scale_fill_viridis(
    name = "Snow",
    discrete = TRUE) +
  theme(plot.title = element_text(size = 12),
        strip.background = element_rect(fill = "black"),
        strip.text = element_text(color = "white", face = "bold"),
        legend.position = "None") 

resp_time_prcp + resp_time_snow
```


## neighborhood
```{r}
library(grid)

neighborhood_freq = 
  incident_dat_2017 %>% 
  na.omit() %>% 
  group_by(neighborhood) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(neighborhood = forcats::fct_reorder(neighborhood, n, 
                                         .asc = TRUE)) %>% 
  ggplot(aes(x = neighborhood, y = n)) + 
  geom_point() +
  labs(x = "neighborhood", y = "Frequency") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45)) 

neighborhood_time = 
  incident_dat_2017 %>%
  na.omit() %>% 
  group_by(neighborhood) %>% 
  summarise(mean_resp_time = mean(response_time)) %>% 
  mutate(neighborhood = forcats::fct_reorder(neighborhood, mean_resp_time, 
                                         .asc = TRUE)) %>% 
  ggplot(aes(x = neighborhood, y = mean_resp_time)) + 
  geom_point() +
  labs(x = "neighborhood", y = "Mean response time") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45))

grid.newpage()
grid.draw(rbind(ggplotGrob(neighborhood_freq), ggplotGrob(neighborhood_time), size = "last"))
```


