---
title: "incident_EDA_yc3242"
author: "Youn Kyeong Chang (uni# yc3242)"
date: "November 12, 2018"
output: html_document
---
```{r}
library(tidyverse)
library(lubridate)
library(xml2)
library(rvest)
library(patchwork)
```

## Response time data - Loaded and tidied data
```{r}
incident_dat_2017 <-  
  read_csv('data/Incidents_Responded_to_by_Fire_Companies.csv',
  col_types = "cicccicciccccccccccciccccccc") %>% 
  janitor::clean_names() %>% 
  #recode date/time
  mutate(incident_date_time = mdy_hms(incident_date_time),
         arrival_date_time = mdy_hms(arrival_date_time)) %>% 
  #select year 2017
  filter(year(incident_date_time) == 2017,
         incident_type_desc == "300 - Rescue, EMS incident, other") %>% 
  select(im_incident_key, incident_date_time, arrival_date_time,
         street_highway:borough_desc) %>% 
  na.omit() %>% 

  # Added response time(minute) variable
  # mutate(response_time = arrival_date_time - incident_date_time) %>%
  mutate(response_time = difftime(arrival_date_time, incident_date_time, 
                                  units = 'secs'),
  # Added hour variable
  hour = hour(incident_date_time),
  date = date(incident_date_time),
  # Added incident_month and incident_day variables from incident_date_time
  incident_date = as.Date(incident_date_time)) %>% 
  separate(incident_date, 
           into = c("incident_year", "incident_month", "incident_day"), 
           sep = "-") %>% 
  select(-incident_year, -street_highway, -borough_desc) %>% 
  mutate(incident_month = as.numeric(incident_month),
         incident_day = as.numeric(incident_day))
```


## Extract ZIP Code Definitions of New York City Neighborhoods
```{r fig.height = 8}
url = "https://www.health.ny.gov/statistics/cancer/registry/appendix/neighborhoods.htm?fbclid=IwAR3N4VlKC1OehRZyEuDYPEAE7AFAEXXIRC11seIBKxA-0fd3g4hL0QvnV20"

xml = read_html(url)

zip_code_table = (xml %>% html_nodes(css = "table")) %>% 
  .[[1]] %>%
  html_table() %>% 
  janitor::clean_names() %>%  
  select(neighborhood, zip_codes) %>% 
  separate(zip_codes, c("a", "b", "c", "d", "e", "f", "g", "h", "i"), 
           sep = ",") %>% 
  gather(key = to_remove, value = zip_code, a:i) %>% 
  select(-to_remove) %>% 
  na.omit() %>% 
  distinct() %>% 
  mutate(zip_code = as.numeric(zip_code))
```

## Add neighborhood variable
```{r}
# To match zip_code to neighborhood table, converted it to numeric
incident_dat_2017 = 
  incident_dat_2017 %>% 
    mutate(zip_code = as.numeric(zip_code)) 

incident_dat_2017 =  
  left_join(incident_dat_2017, zip_code_table, by = "zip_code") 
```

## Response time by area(zip_code)
```{r fig.height = 8}
incident_dat_2017 %>%  
  group_by(zip_code) %>%
  summarise(mean_response_time = mean(response_time, na.rm = TRUE)) %>% 
  mutate(zip_code = forcats::fct_reorder(zip_code, mean_response_time, 
                                         .asc = TRUE)) %>% 
  ggplot(aes(x = zip_code, y = mean_response_time)) + 
  geom_point(size = .5, color = "darkred") +  
  coord_flip() +
  theme_bw()
```


## Response time by area(neighborhood)
```{r fig.height = 8}
incident_dat_2017 %>%
  na.omit(neighborhood) %>% 
  group_by(neighborhood) %>%
  summarise(mean_response_time = mean(response_time, na.rm = TRUE)) %>% 
  mutate(neighborhood = forcats::fct_reorder(neighborhood, mean_response_time, 
                                         .asc = TRUE)) %>% 
  ggplot(aes(x = neighborhood, y = mean_response_time)) + 
  geom_point(size = 1, color = "darkred") +  
  coord_flip() +
  theme_bw()
```


## Importing the weather data
```{r}
library(rnoaa)

nyc_weather_2017 = 
  rnoaa::meteo_pull_monitors("USW00094728", 
                             var = c("PRCP", "TMIN", "TMAX", "SNOW", "SNWD"), 
                             date_min = "2017-01-01", 
                             date_max = "2017-12-31"
                             ) %>% 
  mutate(tmin = tmin/10, tmax = tmax/10, prcp = prcp/10) %>% 
  select(-id)
```


## Combined data with weather data
```{r}
incident_dat_2017 = 
  left_join(incident_dat_2017, nyc_weather_2017, by = "date")
```

## EDA -- mean response time and mean variables from weather data
```{r}
summary_mean = 
incident_dat_2017 %>% 
  group_by(date) %>% 
  summarise(mean_response_time = mean(response_time, na.rm = TRUE),
            mean_prcp = mean(prcp, na.rm = TRUE),
            mean_snow = mean(snow, na.rm = TRUE),
            mean_snwd = mean(snwd, na.rm = TRUE),
            mean_tmax = mean(tmax, na.rm = TRUE),
            mean_tmin = mean(tmin, na.rm = TRUE))
```

## Prcp category
```{r}
nyc_weather_2017 = 
  nyc_weather_2017 %>%
  mutate(prcp_ctg = 
           ifelse(prcp == 0, "NA",
           ifelse((prcp > 0 & prcp <= 25), "low",
           ifelse((prcp > 25 & prcp <= 60), "medium", "high"))))

incident_dat_2017 = 
  inner_join(incident_dat_2017, nyc_weather_2017, by = "date")

incident_dat_2017 %>%
  group_by(date) %>% 
  filter(prcp > 0, 
         neighborhood == "Kingsbridge and Riverdale") %>%
  ggplot(aes(x = prcp_ctg, y = response_time)) +
  geom_violin(aes(fill = factor(prcp_ctg)), alpha = .5) +
  stat_summary(fun.y = mean, geom = "point", size = 4, color = "blue")
```
#
# ```{r}
# summary_mean %>%
#   ggplot(aes(x = mean_snow, y = mean_response_time)) +
#   geom_point() +
#   geom_smooth()
# ```
#
# ```{r}
# summary_mean %>%
#   filter(mean_snwd > 0) %>%
#   ggplot(aes(x = mean_snwd, y = mean_response_time)) +
#   geom_point() +
#   geom_smooth()
# ```
#
# ```{r}
# summary_mean %>%
#   ggplot(aes(x = mean_tmax, y = mean_response_time)) +
#   geom_point() +
#   geom_smooth()
# ```
#
# ```{r}
# summary_mean %>%
#   ggplot(aes(x = mean_tmin, y = mean_response_time)) +
#   geom_point() +
#   geom_smooth()
# ```

## EDA -- yes or no ######
```{r}
incident_time_summary_joined = 
inner_join(incident_dat_2017, summary_mean)

incident_time_summary_joined %>% 
  mutate(prcp = mean_prcp > 0) %>%
  group_by(date) %>% 
  ggplot(aes(x = prcp, y = response_time)) +
  geom_violin(aes(fill = factor(prcp)), alpha = .5) +
  stat_summary(fun.y = mean, geom = "point", size = 4, color = "blue") +
  labs(
    title = "Response time in rainy conditions",
    x = "Precipitation",
    y = "Response time"
  ) +
  viridis::scale_fill_viridis(
    name = "Precipitation",
    discrete = TRUE) +
  theme(plot.title = element_text(size = 12),
        strip.background = element_rect(fill = "black"),
        strip.text = element_text(color = "white", face = "bold"),
        legend.position = "None") 

incident_time_summary_joined %>% 
  mutate(snow = mean_snow > 0) %>% 
  group_by(date) %>% 
  ggplot(aes(x = snow, y = mean_response_time)) +
  geom_violin(aes(fill = factor(snow)), alpha = .5) +
  stat_summary(fun.y = mean, geom = "point", size = 4, color = "blue") +
  labs(
    title = "Mean response time in snowy conditions",
    y = "Mean response time"
  ) +
  viridis::scale_fill_viridis(
    name = "Snow",
    discrete = TRUE) +
  theme(plot.title = element_text(size = 12),
        strip.background = element_rect(fill = "black"),
        strip.text = element_text(color = "white", face = "bold"),
        legend.position = "None") 

resp_time_prcp + resp_time_snow
```

Mean response time seems related to snow but not precipitation. 

# ```{r}
# incident_dat_2017 %>% 
#   group_by(borough_desc) %>% 
#   summarize(mean_response_time = mean(response_time, na.rm = TRUE)) %>% 
#   ggplot(aes(x = borough_desc, y = mean_response_time)) +
#   geom_point()
# ```

## number of incident by zip code vs response time by zip code
```{r}
incident_dat_2017 %>% 
  group_by(zip_code) %>% 
  summarize(n_incident = n(),
            mean_response_time = mean(response_time, na.rm = TRUE)) %>%
  ggplot(aes(x = n_incident, y = mean_response_time)) +
  geom_point() +
  geom_smooth()
```


### High incidence
```{r}
incident_by_date = 
  incident_dat_2017 %>% 
  group_by(zip_code, date) %>% 
  summarize(n_incident = n())  
  
incident_by_zipcode = 
  incident_dat_2017 %>% 
  group_by(zip_code, date) %>% 
  summarize(n_incident = n()) %>% 
  ungroup() %>% 
  group_by(zip_code) %>% 
  summarize(mean_incident = mean(n_incident),
            sd_incident = sd(n_incident),
            std_high_incident = mean_incident + 2*sd_incident,
            std_low_incident = mean_incident - 2*sd_incident)

incident_response_time =
  incident_dat_2017 %>% 
  group_by(zip_code) %>% 
  select(zip_code, response_time, date) 

high_incidence = 
  left_join(incident_by_date, incident_by_zipcode, by = "zip_code") %>% 
  mutate(high_incidence = (n_incident - std_high_incident) > 0) %>% 
  select(zip_code, date, high_incidence)

high_incidence_data = 
  inner_join(incident_dat_2017, high_incidence)

high_incidence_data %>% 
  filter(zip_code == 10001) %>% 
  ggplot(aes(x = high_incidence, y = response_time)) +
  geom_violin(aes(fill = factor(high_incidence)), alpha = .5) +
  stat_summary(fun.y = mean, geom = "point", size = 4, color = "blue") +
  labs(
    title = "Mean response time in rainy conditions",
    x = "high_incidence",
    y = "Mean response time"
  ) +
  viridis::scale_fill_viridis(
    name = "high_incidence",
    discrete = TRUE) +
  theme(plot.title = element_text(size = 12),
        strip.background = element_rect(fill = "black"),
        strip.text = element_text(color = "white", face = "bold"),
        legend.position = "None") 
```

## Number of incidence category
```{r}
incident_by_date = 
  incident_dat_2017 %>% 
  group_by(zip_code, date) %>% 
  summarize(n_incident = n())  
  
incident_by_zipcode = 
  incident_dat_2017 %>% 
  group_by(zip_code, date) %>% 
  summarize(n_incident = n()) %>% 
  ungroup() %>% 
  group_by(zip_code) %>% 
  summarize(mean_incident = mean(n_incident),
            sd_incident = sd(n_incident),
            std_high_incident = mean_incident + sd_incident,
            std_low_incident = mean_incident - sd_incident)

incident_response_time =
  incident_dat_2017 %>% 
  group_by(zip_code) %>% 
  select(zip_code, response_time, date) 

incident_category_data = 
  left_join(incident_by_date, incident_by_zipcode, by = "zip_code") %>% 
  mutate(incident_category = 
           ifelse(((n_incident - std_high_incident) > 0), "high", 
           ifelse(((std_low_incident - n_incident) > 0), "low", "medium"))) %>% 
  select(zip_code, date, incident_category)

incid_cate_joined_data = 
  inner_join(incident_dat_2017, incident_category_data) %>% 
  na.omit() 

# filtered by zip code representative
incid_cate_joined_data %>% 
  filter(zip_code == 10001) %>%  
  ggplot(aes(x = incident_category, y = response_time)) +
  geom_violin(aes(fill = factor(incident_category)), alpha = .5) +
  stat_summary(fun.y = mean, geom = "point", size = 4, color = "blue") +
  labs(
    title = "response time by incident category",
    x = "incident category",
    y = "Mean response time"
  ) +
  viridis::scale_fill_viridis(
    name = "incident category",
    discrete = TRUE) +
  theme(plot.title = element_text(size = 12),
        strip.background = element_rect(fill = "black"),
        strip.text = element_text(color = "white", face = "bold"),
        legend.position = "None") 

# incid_cate_joined_data %>%  
#   group_by(incident_category) %>% 
#   summarize(mean_response_time = mean(response_time)) %>% 
#   ggplot(aes(x = incident_category, y = mean_response_time)) +
#   geom_point()

incid_cate_joined_data %>%  
  group_by(incident_category) %>% 
  ggplot(aes(x = incident_category, y = response_time)) +
  geom_violin(aes(fill = factor(incident_category)), alpha = .5) +
  stat_summary(fun.y = mean, geom = "point", size = 4, color = "blue") +
  scale_y_continuous(limits = c(0,500))
```

## Number of incidence category
```{r}
incident_by_date = 
  incident_dat_2017 %>% 
  group_by(zip_code, date) %>% 
  summarize(n_incident = n())  
  
incident_by_zipcode = 
  incident_dat_2017 %>% 
  group_by(zip_code, date) %>% 
  summarize(n_incident = n()) %>% 
  ungroup() %>% 
  group_by(zip_code) %>% 
  summarize(mean_incident = mean(n_incident))

incident_response_time =
  incident_dat_2017 %>% 
  group_by(zip_code) %>% 
  select(zip_code, response_time, date) 

incident_category_data = 
  left_join(incident_by_date, incident_by_zipcode, by = "zip_code") %>% 
  mutate(incident_category = 
           ifelse(((n_incident - mean_incident) >= 0), "high", "low")) %>%  
  select(zip_code, date, incident_category)

incid_cate_joined_data = 
  inner_join(incident_dat_2017, incident_category_data) %>% 
  na.omit() 

incid_cate_joined_data %>% 
  group_by(incident_category, zip_code) %>% 
  summarize(mean_response_time = mean(response_time, na.rm = TRUE)) %>% 
  ggplot(aes(x = incident_category, y = mean_response_time)) +
  geom_violin(aes(fill = factor(incident_category)), alpha = .5) +
  stat_summary(fun.y = mean, geom = "point", size = 4, color = "blue") +
  labs(
    title = "Mean response time by incidence category",
    x = "incident category",
    y = "Mean response time"
  ) +
  viridis::scale_fill_viridis(
    name = "incident category",
    discrete = TRUE) +
  theme(plot.title = element_text(size = 12),
        strip.background = element_rect(fill = "black"),
        strip.text = element_text(color = "white", face = "bold"),
        legend.position = "None") 
```


## Street closure data - Loaded and tidied data
```{r}
street_closure_2017 <-  
  read_csv('data/Street_Closures_due_to_construction_activities_by_Intersection.csv') %>% 
  janitor::clean_names() %>% 
  #recode date/time
  mutate(work_start_date = mdy_hms(work_start_date),
         work_end_date = mdy_hms(work_end_date),
         work_time = round(difftime(work_end_date, work_start_date, 
                                    units = 'days'), 0)) %>% 
  #select year 2017
  filter(year(work_start_date) == 2017) %>% 
  select(-purpose) %>% 
  na.omit() 
```