---
title: "Weather_EDA"
author: "Abhishek Ajay (aa4266)"
date: "November 11, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rnoaa)
library(tidyverse)

```

Importing the weather data.
```{r data_import, cache = TRUE}
nyc_weather_2017 = 
  rnoaa::meteo_pull_monitors("USW00094728", 
                             var = c("PRCP", "TMIN", "TMAX", "SNOW", "SNWD"), 
                             date_min = "2017-01-01", 
                             date_max = "2017-12-31"
                             )

```
 Data Description: 
 
* id: Weather station ID
* date: Date of observation
* prcp: Precipitation (tenths of mm)
* snow: Snowfall (mm)
* snwd: Snow depth (mm)
* tmax: Maximum temperature (tenths of degrees C)
* tmin: Minimum temperature (tenths of degrees C)

```{r data_tidying}
nyc_weather_2017 = 
  nyc_weather_2017 %>% 
  mutate(tmin = tmin/10, tmax = tmax/10, prcp = prcp/10)
```

top 6 days with **highest temp**
```{r order_1}
nyc_weather_2017 %>% 
  arrange(desc(tmax)) %>% 
  head() %>% 
  knitr::kable()
```

top 6 days with **lowest temp**
```{r order_2}
nyc_weather_2017 %>% 
  arrange(tmin) %>% 
  head() %>% 
  knitr::kable()
```

top 6 days with **highest snow**
```{r order_3}
nyc_weather_2017 %>% 
  arrange(desc(snow)) %>% 
  head() %>% 
  knitr::kable()
```

top 6 days with **least snow**
```{r order_4}
nyc_weather_2017 %>% 
  arrange(snow) %>% 
  head() %>% 
  knitr::kable()
```

top 6 days with **highest snow depth**
```{r order_5}
nyc_weather_2017 %>% 
  arrange(desc(snwd)) %>% 
  head() %>% 
  knitr::kable()
```

top 6 days with **least snow depth**
```{r order_6}
nyc_weather_2017 %>% 
  arrange(snwd) %>% 
  head() %>% 
  knitr::kable()
```
