---
title: "Weather_EDA"
author: "Abhishek Ajay (aa4266)"
date: "November 11, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      fig.width = 6, 
                      fig.asp = .6,
                      out.width = "90%")
library(rnoaa)
library(tidyverse)
library(ggplot2)
library(patchwork)

theme_set(theme_bw() + theme(legend.position = "bottom"))
```

Importing the weather data.
```{r data_import, cache = TRUE}
nyc_weather_2017 = 
  rnoaa::meteo_pull_monitors("USW00094728", 
                             var = c("PRCP", "TMIN", "TMAX", "SNOW", "SNWD"), 
                             date_min = "2017-01-01", 
                             date_max = "2017-12-31"
                             )

```
 Data Description: 
 
* id: Weather station ID
* date: Date of observation
* prcp: Precipitation (tenths of mm)
* snow: Snowfall (mm)
* snwd: Snow depth (mm)
* tmax: Maximum temperature (tenths of degrees C)
* tmin: Minimum temperature (tenths of degrees C)

```{r data_tidying}
nyc_weather_2017 = 
  nyc_weather_2017 %>% 
  mutate(tmin = tmin/10, tmax = tmax/10, prcp = prcp/10)
```

top 6 days with **highest temp**
```{r order_1}
nyc_weather_2017 %>% 
  arrange(desc(tmax)) %>% 
  head() %>% 
  knitr::kable()
```

top 6 days with **lowest temp**
```{r order_2}
nyc_weather_2017 %>% 
  arrange(tmin) %>% 
  head() %>% 
  knitr::kable()
```

top 6 days with **highest snow**
```{r order_3}
nyc_weather_2017 %>% 
  arrange(desc(snow)) %>% 
  head() %>% 
  knitr::kable()
```

top 6 days with **least snow**
```{r order_4}
nyc_weather_2017 %>% 
  arrange(snow) %>% 
  head() %>% 
  knitr::kable()
```

top 6 days with **highest snow depth**
```{r order_5}
nyc_weather_2017 %>% 
  arrange(desc(snwd)) %>% 
  head() %>% 
  knitr::kable()
```

top 6 days with **least snow depth**
```{r order_6}
nyc_weather_2017 %>% 
  arrange(snwd) %>% 
  head() %>% 
  knitr::kable()
```


```{r viz_1}
tmax_plot = 
  nyc_weather_2017 %>% 
  ggplot(aes(x = date, y = tmax)) +
  geom_line() 
  
tmin_plot = 
  nyc_weather_2017 %>% 
  ggplot(aes(x = date, y = tmin)) +
  geom_line()

snow_plot = 
  nyc_weather_2017 %>% 
  ggplot(aes(x = date, y = snow)) +
  geom_line()

snwd_plot = 
  nyc_weather_2017 %>% 
  ggplot(aes(x = date, y = snwd)) +
  geom_line()

prcp_plot = 
  nyc_weather_2017 %>% 
  ggplot(aes(x = date, y = prcp)) +
  geom_line()

((tmax_plot + snow_plot) / (tmin_plot + snwd_plot)) / prcp_plot
```

